---
title: "FRED Tutorial"
author: ""
date: "6/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to use FRED dataset in R

[FRED website](https://fred.stlouisfed.org/) provides various statistics that scholars can utilize in conducting their research. But, it is cumbersome to download data in xls format and import into R for further analysis. So, this tutorial presents a guide on how to import these data easily. With [**fredr**](https://sboysel.github.io/fredr/index.html) package, we can import data into R directly from FRED website.

### Load library

Load libraries for [accessing FRED website](https://sboysel.github.io/fredr/index.html), manipulating data, and visualizing output.

```{r, message=F}
library(fredr)
library(tidyverse)
library(ggplot2)
```

To download data from [FRED website](https://fred.stlouisfed.org/), we need to create an account first, then acquire an API key for accessing [FRED API](https://fred.stlouisfed.org/docs/api/fred/). The API key must be set using **fredr_set_key()** function before using **fredr** function.

```
fredr_set_key("YOUR_API_KEY")
```

### Single series

#### Import data into R directly

**fredr** function retrieve data specified in **series_id**, convert them into table-like format. The range of data are specified in **observation_start** and **observation_end**.

```{r}
dataset <- fredr(
  series_id = "UNRATE",
  observation_start = as.Date("1990-01-01"),
  observation_end = as.Date("2000-01-01")
)
```

#### Inspect data

The data consist of 5 columns and 121 observations.

```{r}
str(dataset)
```

The data are sorted by **date**.

```{r}
dataset %>% head(10)
```

The column **value** indicates unemployment rate. Let's calculate n, max, min, mean, and median of unemployment rate.

```{r}
dataset %>% summarize(n=n(), max=max(value), min=min(value), mean = mean(value), median=median(value))
```

#### Plot data 

Draw line chart of unemployment rate using **ggplot** package.

```{r}
dataset %>%
  ggplot(aes(x=date,y=value)) +
  labs(title="Unemployment rate in USA", x="Year",y="%") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_date(date_breaks = "year" , date_labels = "%Y") +
  geom_line()
```

### Multiple serieses with identical unit

If we compare two series by plotting them in a single chart, use **map_dfr** to import both series simultaneously. For example, we can compare unemployment rate between Korea and USA.

#### Import data

Acronym referring to unemployment rate in Korea is available at [FRED website].(https://fred.stlouisfed.org/series/LRUNTTTTKRM156S)

```{r}
dataset <- map_dfr(c("UNRATE", "LRUNTTTTKRM156S"), fredr)
```

#### Plot data

When we draw line chart, we need to modify the label of legend to reflect the meaning of each series

```{r}
dataset %>%
  ggplot(aes(x = date, y = value, color = series_id)) +
  geom_line(size=1) +
  labs(title="Unemployment rate", x = "Observation Date", y = "Rate", color = "Series") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_discrete(name="Country",labels = c("Korea", "USA"))
```



### Multiple serieses with different unit

#### Import data

In this case, we import multiple series: unemployment rate and job openings.

```{r}
dataset <- map_dfr(c("UNRATE", "JTSJOL"), fredr)
dataset
```

For visualization, we modify table into wider format, where each series become a separate column.

```{r}
dataset_wider <- dataset %>%
  select(date,series_id,value) %>% pivot_wider(names_from="series_id", values_from="value")
dataset_wider
```

Now **UNRATE** and **JTSJOL* became separate column. This conversion is just to make visualization easy.

#### Plot data

Since the unit of **UNRATE** and **JTSJOL** is different, a secondary y axes is added to the right of graph.

```{r}
dataset_wider %>% 
  drop_na() %>%
  ggplot(aes(x=date)) +
  geom_line( aes(y=UNRATE), color="red") + 
  geom_line( aes(y=JTSJOL/1000), color="blue") + 
  scale_y_continuous(
    name = "Unemployment rate",
    sec.axis = sec_axis(~.*1000,name="Job openings")) +
  theme(axis.title.y.left = element_text(colour = "red"),
        axis.title.y.right = element_text(colour = "blue"))
```